---
title: "EDA"
output: html_document
---


## Engine Sizes
by Franklin Doane

```{r setup_packs, echo=FALSE, warning=FALSE, message=FALSE}

options(repos = c(CRAN = "http://cran.rstudio.com/"))


# import code from https://statsandr.com/blog/an-efficient-way-to-install-and-load-r-packages/
if (!("huxtable "%in% rownames(installed.packages()))) {
  install.packages("huxtable", type="binary")
}

if (!("dplyr "%in% rownames(installed.packages()))) {
  install.packages("dplyr", type="binary")
}
```

```{r setup, warning=FALSE, message=FALSE}

library(huxtable)
library(dplyr)

df <- read.csv("bmw.csv")

```

```{r sizes, echo=FALSE}

# remove proceeding model space
df$model <- substr(df$model, 2, nchar(df$model))

# add unit and order
# formatting from https://www.projectpro.io/recipes/convert-integer-float-r
df$engineSizeStr <- paste(noquote(sprintf("%1.1f",df$engineSize)), "L", sep="")

engine_counts <- count(df, engineSizeStr)
barplot(engine_counts$n, cex.names=0.6, names.arg=engine_counts$engineSizeStr)

# table code from https://rfortherestofus.com/2019/11/how-to-make-beautiful-tables-in-r
ht <- count(df, engineSizeStr) %>% 
  as_huxtable(add_colnames=TRUE) %>% 
  set_all_borders(1) %>%
  set_bold(1, everywhere)
t(ht)

```

Here we can see that the 2.0L (4-cylinder) engine and the 3.0 liter (6-cylinder) engine are the most popular

The largest engine size is the 6.6 liter V12, and the smallest is the 0.6 liter 2-cylinder

```{r zeros}


df[df$engineSize == 0 & df$fuelType == "Electric", c("model", "fuelType", "engineSize")]

```

Note that 47 cars have an engine size of 0, and only 2 of them are electric

```{r fuels, echo=FALSE}

ht <- count(df, fuelType) %>% 
  as_huxtable(add_colnames=TRUE) %>% 
  set_all_borders(1) %>%
  set_bold(1, everywhere)
t(ht)

df[df$fuelType == "Electric", c("model", "fuelType", "engineSize")]


```

On the flip side, there are 3 total electric cars, and 1 has a non-0 engine size

Both of these issues may require cleaning


## Models
by Franklin Doane

There are 24 different model types.
The most popular series are 3, 1, 2, 5, and 4

```{r models, echo=FALSE}

ht <- count(df, model) %>% 
  as_huxtable(add_colnames=TRUE) %>% 
  set_all_borders(1) %>%
  set_bold(1, everywhere)
t(ht)

noquote(paste(nrow(table(df$model)), " model types"))

noquote(paste(round(nrow(df[df$model == "3 Series", ]) / nrow(df) * 100, digits=1), "% 3 series"))
noquote(paste(round(nrow(df[df$model == "1 Series", ]) / nrow(df) * 100, digits=1), "% 1 series"))
noquote(paste(round(nrow(df[df$model == "2 Series", ]) / nrow(df) * 100, digits=1), "% 2 series"))
noquote(paste(round(nrow(df[df$model == "5 Series", ]) / nrow(df) * 100, digits=1), "% 5 series"))
noquote(paste(round(nrow(df[df$model == "4 Series", ]) / nrow(df) * 100, digits=1), "% 4 series"))


```

Also worth noting is that all series names have proceeding spaces which should be cleaned

```{r pies, echo=FALSE, warning=FALSE, message=FALSE}

counts <- count(df, model)
counts <- counts[order(counts$n), ]

is_suv <- substr(counts$model, 1, 1) == "X"
suvs <- counts[is_suv, ]
counts_simplified <- counts[!is_suv, ]
counts_simplified <- rbind(counts_simplified, list("X models", sum(suvs$n)))
# simplified away all suvs into a single category

is_z <- substr(counts_simplified$model, 1, 1) == "Z"
zs <- counts_simplified[is_z, ]
counts_simplified <- counts_simplified[!is_z, ]
counts_simplified <- rbind(counts_simplified, list("Z models", sum(zs$n)))
# simplified away all Zs into a single category

is_m <- substr(counts_simplified$model, 1, 1) == "M"
ms <- counts_simplified[is_m, ]
counts_simplified <- counts_simplified[!is_m, ]
counts_simplified <- rbind(counts_simplified, list("M models", sum(ms$n)))
# simplified away all Ms into a single category

is_i <- substr(counts_simplified$model, 1, 1) == "i"
is <- counts_simplified[is_i, ]
counts_simplified <- counts_simplified[!is_i, ]
counts_simplified <- rbind(counts_simplified, list("electric", sum(is$n)))
# simplified away all Is into a single category

pie(counts$n, counts$model, radius=1.075, cex=0.6)
pie(counts_simplified$n, counts_simplified$model, radius=1.075, cex=0.6)


```

The second pie chart is simplified a bit, grouping some of the model types

X models: suvs

M models: high performance

Z models: 2-seat roadsters


## Transmission
by Will Ward

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```

```{r}
library(tidyverse)


```

```{r}
df <- read.csv("bmw.csv")
summary(df)
```

```{r}
hist(df$price, breaks=100)

transmission_A <- df %>% filter(transmission=="Automatic")
transmission_Semi <- df %>% filter(transmission=="Semi-Auto")
transmission_M <- df %>% filter(transmission=="Manual")

hist(transmission_A$price, breaks = 100, xlim=c(0,75000))
hist(transmission_Semi$price, breaks = 100, xlim=c(0,75000))
hist(transmission_M$price, breaks = 100, xlim=c(0,75000))


```

```{r}
anova_transmisison <- aov(price ~ transmission, data = df)
summary(anova_transmisison)
# transmission Sum sq = variation explained by transmission type
# residuals Sum sq = variation unexplained by transmission type

# n^2 = t-SumSq/r-SumSq = 0.189
# around 19% of the variance in price is explained by transmission

```

```{r}
TukeyHSD(anova_transmisison)
#                          diff       lwr       upr p adj
# Manual-Automatic    -7781.573 -8407.339 -7155.807     0
# Semi-Auto-Automatic  4940.724  4405.696  5475.752     0
# Semi-Auto-Manual    12722.297 12127.152 13317.442     0

# manual cars are about $7,782 cheaper than automatics
# Semi-Auto cars are about $4,941 more expensive than automatics
# Semi-Auto cars are about $12,722 more expensive than manuals
```

```{r}
colnames(df)
```

```{r}
anova_model <- aov(price ~ model, data=df)
summary(anova_model)

# 6.900e11/(6.900e11 + 7.148e11) = .49 | 49% of variation in price is explained by model
```

```{r}
model_transmission <- lm(price~transmission, data=df)
summary(model_transmission)

model_carM <- lm(price~model, data=df)
summary(model_carM)
```

```{r}
model_MT <- lm(price ~ model + transmission, data=df)
summary(model_MT)

# multiple r^2 value = .572 | 57% of variation in price explained by model and transmission. #notice that .49 + .19 is not 0.572. Transmission adds some unique predictive power on top of model, but its explainability overlaps with model somewhat. 
```


## Year and Mileage
by Slate Lee

```{r ggplot library}
library(ggplot2)
```

```{r summary of data}
bmw_data <- read.csv("bmw.csv", header = TRUE, sep = ",")
summary(bmw_data)
```
Notably, this data set has 10781 entries of cars from year 1996 - 2020 and with a mileage from 1 - 214000.

```{r plot}
ggplot(data = bmw_data, aes(x = year, y = mileage)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "blue")
```

A basic plot of year and mileage shows a negative correlation: when year increases the mileage goes down. This is expected and obvious of real data, since newer cars are less used. Adding a linear model ("lm") shows a diagonal line. However a curved line ("loess") better fits the the data especially when considering the older cars, since their usage also seems to fall off.

```{r correlation coeffecient}
correlation_coeff <- cor(bmw_data$year, bmw_data$mileage)
print(paste("Correlation:", round(correlation_coeff, 3)))
```

When correlation is calculated the resulting value is "-0.774," again showing the expected negative relationship.

```{r mileage-frequency}
mileage_counts <- sort(table(bmw_data$mileage), decreasing = TRUE)
head(mileage_counts, 100)
```

The above table counts the amount of times any given mileage value occurs, and has them ordered from highest to lowest. Many of the most common mileage values are rounded to the nearest hundred or thousand (e.g. "1000", "5000", "100"). This shows that the original data set has been rounded, and are not the sort of random values which we would expect. The table does show a mode of 123 which occurred 151 times. This is a strangely specific value especially given that the surrounding numbers 122 and 124 do not show up on the top 100 values. This is evidence that the may not accurately correspond to real world data, or has been adjusted.


## Fuel Type
by Shaye Shankar

``` {r}
library(ggplot2)
```

``` {r}
df <- read.csv("bmw.csv")
summary(df)
```

``` {r count of vehicles}
ggplot(df, aes(x=fuelType)) +
  geom_bar() +
  labs(title="Count of Vehicles by Fuel Type")
```
``` {r}
fuel_counts_base <- table(df$fuelType)
print(fuel_counts_base)
```
There's the most numerical and consistent data on petrol cars and diesel cars. There's only 3 instances for electric cars. The 36 "other" fuelType category may be the result of missing data of other cars.


## MPG
by Shaye Shankar

``` {r}
ggplot(df, aes(x=fuelType, y=mpg, fill=fuelType)) +
  geom_violin(trim=FALSE) +
  geom_boxplot(width=0.1, fill="white") +
  labs(title="MPG Distribution by Fuel Type")
```

Petrol and Diesel cars are the only real trustworthy information for looking at miles and the data makes seems real as diesel, on average, has just slightly more mpg than petrol. Electric has the highest range and likely should not be included on this plot, since it is an electric car and does not have a real mileage per gallon. There are a lot of inconsistencies with MPG for Hybrids, with among the highest being 470 miles per gallon, which is very likely an error.

``` {r MPG vs Engine Size by Fuel Type}
ggplot(df, aes(x=mpg, y=engineSize)) +
  geom_point() +
  facet_wrap(~ fuelType) +
  labs(title="MPG vs Engine Size by Fuel Type")
```

Electric vehicles do not have an engine size, and thus should not be included in the comparison.There appears to be a weaker correlation between engine size and mpg for hybrids and a massive outlier with 400+ mpg. Diesel and petrol have the most consistent data, following a trend of having higher miles per gallon the smaller the engine size is, unless the engine size is 0 in which case the data is incorrect as this should be impossible.


