---
title: "Franklin's EDA"
output: html_document
date: "2025-11-10"
---

```{r setup_packs, echo=FALSE, warning=FALSE, message=FALSE}

options(repos = c(CRAN = "http://cran.rstudio.com/"))


# import code from https://statsandr.com/blog/an-efficient-way-to-install-and-load-r-packages/
if (!("huxtable "%in% rownames(installed.packages()))) {
  install.packages("huxtable", type="binary")
}

if (!("dplyr "%in% rownames(installed.packages()))) {
  install.packages("dplyr", type="binary")
}
```

```{r setup, warning=FALSE, message=FALSE}

library(huxtable)
library(dplyr)

df <- read.csv("bmw.csv")

```

## Engine Sizes


```{r sizes, echo=FALSE}

# remove proceeding model space
df$model <- substr(df$model, 2, nchar(df$model))

# add unit and order
# formatting from https://www.projectpro.io/recipes/convert-integer-float-r
df$engineSizeStr <- paste(noquote(sprintf("%1.1f",df$engineSize)), "L", sep="")

engine_counts <- count(df, engineSizeStr)
barplot(engine_counts$n, cex.names=0.6, names.arg=engine_counts$engineSizeStr)

# table code from https://rfortherestofus.com/2019/11/how-to-make-beautiful-tables-in-r
ht <- count(df, engineSizeStr) %>% 
  as_huxtable(add_colnames=TRUE) %>% 
  set_all_borders(1) %>%
  set_bold(1, everywhere)
t(ht)

```

Here we can see that the 2.0L (4-cylinder) engine and the 3.0 liter (6-cylinder) engine are the most popular

The largest engine size is the 6.6 liter V12, and the smallest is the 0.6 liter 2-cylinder

```{r zeros}


df[df$engineSize == 0 & df$fuelType == "Electric", c("model", "fuelType", "engineSize")]

```

Note that 47 cars have an engine size of 0, and only 2 of them are electric

```{r fuels, echo=FALSE}

ht <- count(df, fuelType) %>% 
  as_huxtable(add_colnames=TRUE) %>% 
  set_all_borders(1) %>%
  set_bold(1, everywhere)
t(ht)

df[df$fuelType == "Electric", c("model", "fuelType", "engineSize")]


```

On the flip side, there are 3 total electric cars, and 1 has a non-0 engine size

Both of these issues may require cleaning


## Models

There are 24 different model types.
The most popular series are 3, 1, 2, 5, and 4

```{r models, echo=FALSE}

ht <- count(df, model) %>% 
  as_huxtable(add_colnames=TRUE) %>% 
  set_all_borders(1) %>%
  set_bold(1, everywhere)
t(ht)

noquote(paste(nrow(table(df$model)), " model types"))

noquote(paste(round(nrow(df[df$model == "3 Series", ]) / nrow(df) * 100, digits=1), "% 3 series"))
noquote(paste(round(nrow(df[df$model == "1 Series", ]) / nrow(df) * 100, digits=1), "% 1 series"))
noquote(paste(round(nrow(df[df$model == "2 Series", ]) / nrow(df) * 100, digits=1), "% 2 series"))
noquote(paste(round(nrow(df[df$model == "5 Series", ]) / nrow(df) * 100, digits=1), "% 5 series"))
noquote(paste(round(nrow(df[df$model == "4 Series", ]) / nrow(df) * 100, digits=1), "% 4 series"))


```

Also worth noting is that all series names have proceeding spaces which should be cleaned

```{r pies, echo=FALSE, warning=FALSE, message=FALSE}

counts <- count(df, model)
counts <- counts[order(counts$n), ]

is_suv <- substr(counts$model, 1, 1) == "X"
suvs <- counts[is_suv, ]
counts_simplified <- counts[!is_suv, ]
counts_simplified <- rbind(counts_simplified, list("X models", sum(suvs$n)))
# simplified away all suvs into a single category

is_z <- substr(counts_simplified$model, 1, 1) == "Z"
zs <- counts_simplified[is_z, ]
counts_simplified <- counts_simplified[!is_z, ]
counts_simplified <- rbind(counts_simplified, list("Z models", sum(zs$n)))
# simplified away all Zs into a single category

is_m <- substr(counts_simplified$model, 1, 1) == "M"
ms <- counts_simplified[is_m, ]
counts_simplified <- counts_simplified[!is_m, ]
counts_simplified <- rbind(counts_simplified, list("M models", sum(ms$n)))
# simplified away all Ms into a single category

is_i <- substr(counts_simplified$model, 1, 1) == "i"
is <- counts_simplified[is_i, ]
counts_simplified <- counts_simplified[!is_i, ]
counts_simplified <- rbind(counts_simplified, list("electric", sum(is$n)))
# simplified away all Is into a single category

pie(counts$n, counts$model, radius=1.075, cex=0.6)
pie(counts_simplified$n, counts_simplified$model, radius=1.075, cex=0.6)


```

The second pie chart is simplified a bit, grouping some of the model types

X models: suvs

M models: high performance

Z models: 2-seat roadsters



